Spike-and-Slab LASSO Generalized Additive Models and Scalable Algorithms for High-Dimensional Data Analysis

### 2022-04-02 Manuscript/R&R/00-response_letter.Rmd

\bigskip

# Response to Associate Editor
_Comments to the authors:_

_Please follow carefully the comments, criticisms and suggestions of the three referees._

__Response__: We summarize the major changes in our manuscript as a response to address the comments from the referees:
\begin{itemize}
    \item According to Reviewer \#2's and Reviewer \#4's comments, we modified the prior distribution of the non-linear effects to reflect the effect hierarchy principle for bi-level selection. We updated the simulation and real data analysis results.
    \item We streamlined the structure of the manuscript, moving the previous EM-IWLS fitting algorithm subsection to the supporting information for three reasons. First of all, moving can help the audience focus on the model formulation and EM-Coordinate Descent algorithm, avoiding confusion and distraction. Secondly, many useful comments raised by the reviewers on the EM-IWLS algorithm, e.g. constructing hypothesis testings and confidence band, require a more dedicated discussion. We are preparing another manuscript discussing those. Lastly, with the added variable selection performance, the manuscript would be too lengthy if keeping the EM-IWLS in the main body. The readability of the manuscript would be damaged. Relocating the EM-IWLS section solves the dilemma without hurting the integrity of the current manuscript that emphasizes scalability.
    \item According to Reviewer \#1's and Reviewer \#4's comments, we added the linear LASSO model and spikeSlabGAM model among the models to be compared in the simulation study. In addition, we added a new set of simulation scenarios where the functions of predictors are linear.  
    \item According to Reviewer \#4's comment, we summarized the variable selection and bi-level selection performance in the simulation study section. The computation performance of spikeSlabGAM is recorded and addressed too.
    \item According to Reviewer \#4's comment, we added the plot showing the estimated functions of predictors for real-data analysis.
    \item We updated the feature screening procedure in our real-data analysis of the Emory Cardiovascular Biobank dataset. Our previous screening method, fitting univariate GAM, made the "double-dipping" mistake that would overestimate the prediction performance of any downstream prediction model. We updated the screening procedure with variance-based selection, and the prediction of both methods dropped. Nevertheless, the proposed model still outperforms the compared method. See change on P9.
    \item We fixed the additional typos and grammar errors according to Review \#2. The revision of the manuscript is marked in \bg{red} (roughly due to the technical difficulties associated with Latex). Deletions are not highlighted.
\end{itemize}

### 2022-04-02 Manuscript/R&R/01-reviewer_1.Rmd
correct typo

__Response__: Thank you for the comment. We added linear LASSO models to our simulation studies and see that the proposed model predicts as good as linear LASSO models when the underlying signals are linear, and substantially better when the underlying signals are nonlinear. This empirically demonstrates the proposed model is flexible and does not need additional metrics to decide if the model is appropriate to model linear or non-linear signals. The simulation results also match with our expectation. The proposed model, by design, can automatically detect if the underlying is linear or nonlinear, framed as bi-level selection problem in the manuscript. When the underlying signal is linear, the nonlinear component of the smoothing function would not be included in the model due to the proposed two-part spike-and-slab lasso prior.

@@

__Response__: Thank you for the comment. The purpose of mentioning the other several priors in the previous manuscript is to describe the generalization of the proposed algorithms. In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalable model. We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily generalize to other family of priors or other family of smoothing functions if desired.").

Compared to other priors, the spike-and-slab double exponential prior provides three advantages. First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients. Hence, the smoothing functions can be estimated more accurately. Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straight forward. Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.

@@

__Response__: We thank the review for the comments. For prediction purpose, the proposed model can be flexibly used in the high-dimensional data analysis without any checking or testing. The two-part spike-and-slab LASSO prior automatically detects if the underlying signal is linear or complex. When the signal is linear, the smoothing function of the predictor only includes the linear part; when the signal is complex, cubic spline (implemented in the package) is used to estimate the complex signal. As a supporting evidence, our added simulation demonstrates the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see response to the Comment 1).

@@

In conclusion, the proposed model do not addition detection for prediction purpose, and have different utility for bi-level selection.

>>>>

__Response__: Thank you for the comment. We added linear LASSO models to our simulation studies and see that the proposed model predicts as good as linear LASSO models when the underlying signals are linear, and substantially better when the underlying signals are nonlinear. This empirically demonstrates that the proposed model is flexible and does not need additional metrics to decide if the model is appropriate to model linear or non-linear signals. The simulation results also match our expectations. The proposed model, by design, can automatically detect if the underlying is linear or nonlinear, framed as the bi-level selection problem in the manuscript. When the underlying signal is linear, the nonlinear component of the smoothing function would not be included in the model due to the proposed two-part spike-and-slab lasso prior.

@@

__Response__: Thank you for the comment. The purpose of mentioning the other several priors in the previous manuscript is to describe the generalization of the proposed algorithms. In the updated manuscript, we move the subsection describing the other priors to the supporting information so that the readers can focus on understanding the proposed scalable model. We included an additional sentence in the discussion section (see P10 "Moreover, the proposed model can easily generalize to other families of priors and smoothing functions if desired.").

Compared to other priors, the spike-and-slab double exponential prior provides three advantages. First of all, the spike-and-slab double exponential prior provides a locally adaptive shrinkage when estimating the coefficients. Hence, the smoothing functions can be estimated more accurately. Secondly, the spike-and-slab exponential prior encourages a sparse solution, making variable selection straightforward. Thirdly, the spike-and-slab double exponential prior motivates a scalable algorithm, the EM-CD algorithm, for model fitting, and hence is more feasible for high-dimensional data analysis.

@@

__Response__: We thank the reviewer for the comments. For prediction purposes, the proposed model can be flexibly used in the high-dimensional data analysis without any checking or testing. The two-part spike-and-slab LASSO prior automatically detects if the underlying signal is linear or complex. When the signal is linear, the smoothing function of the predictor only includes the linear part; when the signal is complex, cubic splines (implemented in the package) are used to estimate the complex signal. As supporting evidence, our added simulation demonstrates the proposed model predicts as good as, if not better than, the linear LASSO model, even when signals are linear and sparse (see the response to Comment 1).

@@

In conclusion, the proposed model does not require addition detection when used for prediction, and has different utility for bi-level selection.

### 2022-04-02 Manuscript/R&R/03-reviewer_4.Rmd
correct typo

__Response__: We thank the reviewer for the comment. We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously. This prior formulation is motivated by effect hierarchy discussed in Chipman (2004). Now, the inclusion of the nonlinear effects is strictly dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection. The modification of the prior doesn't complicate the model fitting algorithm. The $\gamma$ can be analytically integrated out of the density function and requires minimum change of the proposed algorithm. In other words, the proposed model remain feasible for high-dimensional data analysis and retains the claimed computational advantages.
\bigskip

@@

__Response__: We thank the review for the comments. We made change to our simulation study section, mainly adding variable selection and a new set of simulations that the predictors are all linear. The suggested model spikeSlabGAM and linear Lasso model as the benchmark were added to the models of comparison. When reporting the variable selection performance, we considered three metrics, positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient. Bi-level selection performance was provided for the two applicable models, the proposed model and spikeSlabGAM. Please see the updated simulation study on P7-9.

@@

__Response__: We thank the review for the comment. In the update manuscript, we add the suggested model, spikeSlabGAM, in the simulation study, where all the performance metrics and computation time are recorded and compared. We see that the proposed method still preserves the computation advantage over spikeSlabGAM, particularly in the high-dimensional cases.
\bigskip

@@

__Response__: We thank the reviewer for the comment. We agree that using "nonparametric methods" to differentiate from "black box" methods would be misleading. However, we respectively argue that our word choice, i.e. "nonparametric regression models", is not misleading. First of all The black box models normally are not treated as _regression_ models. Moreover, the term "nonparametric regression models" has been consistently used in the generalized additive models literature, for example @Hastie1987. Last but not least, to the authors' knowledge, there is a subtle distinction between nonparametric regression models and semiparametric regression models: semiparametric regression models assume some predictors have linear effects, while nonparametric regression models assume unknown functions for all predictors. To give an example, see @du2010. The subtle definition distinction would not matter as much in this manuscript, but would be relevant to the future work, incorporating predictor structure in the additive model, e.g. gene network. We hope to establish the consistency in the word choice to avoid future confusion.

>>>>

__Response__: We thank the reviewer for the comment. We changed the prior distribution of the non-linear effects inclusion indicator to be $\gamma^\tp|\gamma, \theta \sim Bin(1, \gamma\theta)$, in comparison to $\gamma^\tp| \theta \sim Bin(1, \theta)$ previously. This prior formulation is motivated by the effect hierarchy discussed in Chipman (2004). Now, the inclusion of the nonlinear effects is strictly dependent on the inclusion of the linear effects, and hence, reflects the hierarchy of bi-level selection. The modification of the prior doesn't complicate the model fitting algorithm. The $\gamma$ can be analytically integrated out of the density function and requires minimum change of the proposed algorithm. In other words, the proposed model remains feasible for high-dimensional data analysis and retains the claimed computational advantages.
\bigskip

@@

__Response__: We thank the reviewer for the comments. We made changes to our simulation study section, mainly adding variable selection and a new set of simulations that the predictors are all linear. The suggested model spikeSlabGAM and linear Lasso model as the benchmark were added to the models of comparison. When reporting the variable selection performance, we considered three metrics, positive predictive value (precision), true positive rate (recall), and Matthews correlation coefficient. Bi-level selection performance was provided for the two applicable models, the proposed model and spikeSlabGAM. Please see the updated simulation study on P7-9.

@@

__Response__: We thank the reviewer for the comment. In the update manuscript, we add the suggested model, spikeSlabGAM, in the simulation study, where all the performance metrics and computation time are recorded and compared. We see that the proposed method still preserves the computation advantage over spikeSlabGAM, particularly in the high-dimensional cases.
\bigskip

@@

__Response__: We thank the reviewer for the comment. We agree that using "nonparametric methods" to differentiate from "black box" methods would be misleading. However, we respectively argue that our word choice, i.e. "nonparametric regression models", is not misleading. First of all The black box models normally are not treated as _regression_ models. Moreover, the term "nonparametric regression models" has been consistently used in the generalized additive models literature, for example @Hastie1987. Last but not least, to the authors' knowledge, there is a subtle distinction between nonparametric regression models and semiparametric regression models: semiparametric regression models assume some predictors have linear effects, while nonparametric regression models assume unknown functions for all predictors. To give an example, see @du2010. The subtle definition distinction would not matter as much in this manuscript, but would be relevant to the future work, incorporating predictor structure in the additive model, e.g. gene network. We hope to establish consistency in the word choice to avoid future confusion.
